---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "figures/",
  out.width = "100%",
  fig.retina = 2,
  echo = FALSE
)
```


# Visualizing the Corona (COVID-19) pandemic

## Basic idea

I generally think that the news do a good job in describing and visualizing the corona pandemic. However, there are two things that I believe are quite problematic and that I am missing the any news coverage on the pandemic. 

1. Visualizations of the growth curves often log-transform the x-axis instead of showing the actual *exponential* growth.
2. Visualizations almost never plot *total* cases against a) deaths, b) recovered, AND c) active cases. 

With regard to the former, I recently wrote [this blogpost](http://philippmasur.de/blog/2020/03/13/understanding-exponential-growth-the-corona-pandemic/) that explains why exponential growth is so hard to grasp. With regard to the latter, I believe that visualizations would become a bit more informative if we plot total cases, active cases, recovered cases, and deaths. If we plot all *four* curves simultaneously, patterns emerge that might tell us something about the "phase"" that a country is it. We may also see whether a country succeeds in stopping the infections (BUT: see disclaimer further below!). In what follows, I am using the data on worldwide total infections, death rates, and number of recoveries to produce these (potentially) a bit more insightful visualizations. 

#### Where does the data come frome?

The analyses and visualizations are based on the data provided by the John Hopkins University in the [Official 2019 Novel Coronavirus COVID-19 (2019-nCoV) Data Repository](https://github.com/CSSEGISandData/COVID-19). The same data sets are used to constantly update this visual dashboard: https://coronavirus.jhu.edu/map.html

**IMPORTANT DISCLAIMER:**
Although I do believe that these data help us to understand the pandemic, they are nonetheless very imprecise. When we want to make sense of positive test results (i.e., total cases and in the long run mortality rates and recovering processes), we need to know how many tests were conducted. South Korea, for example, has tested > 270,000 people, which amounts to more than 5,200 tests per million inhabitants. The numbers in South Korea are hence more trustworthy than the numbers of all other countries. A high diagnostic capacity at scale is hence key to epidemic control as it provides us with precise estimates and growth rate predictions (for more information on this, see [this article](https://www.sciencemag.org/news/2020/03/coronavirus-cases-have-dropped-sharply-south-korea-whats-secret-its-success?fbclid=IwAR3BnhqQMxCdu8-fQelEkWIDQn-j9UASV773Xl-WbIy8l7M5ZVSQpHFgkL8) in Science). For this reasons, I believe that the value of these visualization lies NOT in comparing the actual numbers, but in understanding the patterns that emerge by comparing all four curves. 

Furthermore, I would like to emphasize that I am not an expert on epidemology or virus outbreaks and I am not working in the health sector. On this page, I am only visualizing the data by the John Hopkins University in a different way that most news or other outlets do. Reliance on the these visualizations for medical guidance or use of these visualization in commerce is strictly prohibited.

#### Will these figures be updated?

Yes, I will update these figures every morning. The last update was made on `r Sys.time()`. The data of the John Hopkins University, however, are always updated at 23:59. What you see is hence the situation on `r Sys.Date()-1` at 23:59:00. Also bear in mind that the reporting of cases is somewhat delayed so that it is very likely that the actual numbers are higher.

## Visualizations

If you are interested in the R code, please see the [README.rmd](https://github.com/masurp/corona_analyses/blob/master/README.rmd). 

```{r, message = F, warning = F}
library(tidycovid19)
library(tidyverse)
library(lubridate)

# Function to plot figures
plot_graph <- function(data, size = .75){
  ggplot(data, aes(x = date, y = value, color = key)) +
  geom_line(size = size) +
  scale_color_brewer(palette = "Dark2") +
  facet_wrap(~country, scale = "free_y") +
  theme_linedraw() +
  labs(x = "date", y = "cases", color = "")
}

# Get & wrangle data ####
data <- download_jhu_csse_covid19_data() 

data <- data %>%
  select(-iso3c, -timestamp) %>%
  mutate(active = confirmed - (deaths + recovered))

```

### 1. Analyzing China, South Korea, Italy, and Germany

In a first step, I am comparing China, South Korea, Italy, and Germany. Why these countries? These four countries are at very different stages. China was the first to experience the outbreak and they have almost contained the spreading of the virus by now. South Korea is starting to control the virus outbreak. Italy is currently experiencing one of the worst outbreaks and drastic measures have been taken in the last weeks. The virus has reached Germany considerably later, but the growth rate is very steep at this moment. 

```{r example,results = 'asis'}
# Current cases 
table <- data %>%
  filter(date == Sys.Date()-1) %>%
  filter(country == "China" | country == "Germany" | 
         country == "Italy" | country == "Korea, South") %>%
  arrange(desc(confirmed))
papaja::apa_table(table, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```  

```{r, fig.height=6, fig.width=7}
# Example plot for China, Germany, Italy and South Korea
data %>%
  filter(confirmed >= 1) %>%
  filter(country == "China" | country == "Germany" | 
         country == "Italy" | country == "Korea, South") %>%
  mutate(country = factor(country, 
                          levels = c("China", "Korea, South", 
                                     "Italy", "Germany"))) %>%
  gather(key, value, -country, -date) %>%
  plot_graph

ggsave("figures/plot_1.png",
       width = 7,
       height = 6)
```


### 2. Worldwide developments

#### Europe

```{r europe, results = 'asis', comment = ""}
europe <- c("Italy", "Spain", "Germany", "France", 
            "Switzerland", "United Kingdom", "Netherlands", 
            "Norway", "Austria", "Sweden", "Belgium", "Denmark",
            "Portugal", "Czechia", "Ireland", "Turkey")

# Current cases 
table3 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country %in% europe) %>%
  arrange(desc(confirmed))

# Plot table
papaja::apa_table(table3, 
                  format = "html", 
                  digits = 0, 
                  align = c("l", "c", rep("r", 4)))
``` 

```{r europe_plot, fig.height=5.75, fig.width = 9}
data %>%
  filter(confirmed >= 1) %>%
  filter(country %in% europe) %>%
  gather(key, value, -country, -date) %>%
    mutate(date = lubridate::ymd(date)) %>%
  plot_graph(size = .6) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggsave("figures/plot_3.png",
       width = 9,
       height = 5.75)
```

#### North, Middle and South America

```{r,results = 'asis'}
america <- c("US", "Canada", "Brazil")

# Current cases 
table4 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country %in% america) %>%
  arrange(desc(confirmed))
papaja::apa_table(table4, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```  

```{r northamerica, fig.height=3, fig.width = 9}
data %>%
  filter(confirmed >= 1) %>%
  filter(country %in% america) %>%
  gather(key, value, -country, -date) %>%
  mutate(date = lubridate::ymd(date)) %>%
  plot_graph

ggsave("figures/plot_4.png",
       width = 9,
       height = 3)
```

#### Middle East

```{r,results = 'asis'}
# Current cases 
table5 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country %in% c("Iran", "Qatar", 
                        "Israel", "Pakistan")) %>%
  arrange(desc(confirmed))
papaja::apa_table(table5, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```  

```{r middleeast, fig.height=7, fig.width = 8}
data %>%
  filter(confirmed >=1) %>%
  filter(country %in% c("Iran", "Qatar", 
                        "Israel", "Pakistan")) %>%
  gather(key, value, -country, -date) %>%
  mutate(date = lubridate::ymd(date)) %>%
  plot_graph

ggsave("figures/plot_5.png",
       width = 8,
       height = 7)
```

#### Asia

```{r,results = 'asis'}
# Current cases 
table6 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country %in% c("China", "Korea, South", "Japan", "Malaysia")) %>%
  arrange(desc(confirmed))
papaja::apa_table(table6, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```

```{r asia, fig.height=7, fig.width = 8}
data %>%
  filter(confirmed >= 1) %>%
  filter(country %in% c("China", "Korea, South", "Japan", "Malaysia")) %>%
  gather(key, value, -country, -date) %>%
  mutate(date = lubridate::ymd(date)) %>%
  plot_graph

ggsave("figures/plot_6.png",
       width = 8,
       height = 7)
```


### 3. Alternative visualizations

One thing that is constantly debatted is how to visualize growth at all. Log-transform the y-axis or not? Plot against the date? Plot against days after 100th case? Plot something entirely different?

Here, I would like to explain some differences between visualizations that have been used in the media. All of them are helpful in their own regard. 

1. LEFT: Here, I plotted total cases against days after the 100th cases *without* logarithmizing the y-axis. 

2. MIDDLE: The y-axis is logarithmized. This figure is often shown in the news. 

3. RIGHT: New cases per 7 days (y-axis) are plotted against total cases (x-axis), both axes are logarithmized.

We see clearly that each plot has benefits and weaknesses. The first example provides perhaps the best comparison of the total numbers and shows quite drastically how exponential growth looks likes. S-curves represent positive developments towars a slower growth. Yet, changes are barely identifiable. The second example makes the actual growth more comparable and by logarithmizing the y-axsis, we can actually see changes in the growth. Here, it seesm that most countries are actually starting to slow the growth. Finally, the last example is hard to understand generally, but it shows best whether some sort of measure is working. The curve needs to sink drastically, otherwise, the growth is continuing uncontrollibly. 

```{r warning = F, message = F, fig.height=5, fig.width = 12}

pg1 <- data %>%
  filter(confirmed >= 100) %>%
  filter(country == "Germany" | country == "US" |
         country == "Italy" | country == "Korea, South") %>%
  group_by(country) %>%
  mutate(days = 1:n()) %>%
  ggplot(aes(x = days, y = confirmed, color = country)) + 
  geom_line(size = .75) +
  theme_linedraw() +
  scale_color_brewer(palette = "Dark2") +
  theme(legend.position = "none") +
  labs(x = "days after 100 cases", y = "total cases", color = "")

pg2 <- data %>%
  filter(confirmed >= 100) %>%
  filter(country == "Germany" | country == "US" |
         country == "Italy" | country == "Korea, South") %>%
  group_by(country) %>%
  mutate(days = 1:n()) %>%
  ggplot(aes(x = days, y = confirmed, color = country)) + 
  geom_line(size = .75) +
  theme_linedraw() +
  scale_y_log10() +
  scale_color_brewer(palette = "Dark2") +
  theme(legend.position = "bottom") +
  labs(x = "days after 100 cases", y = "total cases (log)", color = "")

pg3 <- data %>%
  filter(confirmed >= 100) %>%
  filter(country == "Germany" | country == "US" |
         country == "Italy" | country == "Korea, South") %>%
  group_by(country) %>%
  mutate(new_cases = confirmed - lag(confirmed, 7)) %>%
  ungroup %>%
  ggplot(aes(x = confirmed, y = new_cases, color = country)) + 
  geom_line(size = .75) +
  scale_x_log10() +
  scale_y_log10() +
  theme_linedraw() +
  scale_color_brewer(palette = "Dark2") +
  theme(legend.position = "none") +
  labs(x = "total cases (log)", y = "new cases (log)", color = "")

cowplot::plot_grid(pg1, pg2, pg3,
                   align = "h",
                   axis = "lbr",
                   ncol = 3)
  
```


