---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/",
  out.width = "100%",
  fig.retina = 2
)
```


# Visualizing the Corona (COVID-19) pandemic

## Basic idea

I generally think that the news do a good job in describing and visualizing the corona pandemic. However, there are two things that I believe are quite problematic and that I am missing the any news coverage on the pandemic. 

1. Visualizations of the growth curves often log-transform the x-axis instead of showing the actual *exponential* growth.
2. Visualizations almost never plot *total* cases against a) deaths, b) recovered, AND c) active cases. 

With regard to the former, I recently wrote [this blogpost](http://philippmasur.de/blog/2020/03/13/understanding-exponential-growth-the-corona-pandemic/) that explains why exponential growth is so hard to grasp.
With regard to the latter, I believe that showing only the total cases growth curve can be misleading or at least is not sufficient to understand the pandemic and to try to judge whether certain measures work or not. From my point of view, these visualizations would become a bit more informative, if we also look at the active cases vs. recovered cases vs. deaths. Looking at all *four* curves simultaneously, we can at least partly see how well a country succeeds in stopping the infections (BUT: see disclaimer further below!)

In what follows, I am using actual data on worldwide total infections, death rates, and number of recoveries to produce these more insightful visualizations. 

### Where does the data come frome?

The analyses and visualizations are based on the data provided by the John Hopkins University in the [Official 2019 Novel Coronavirus COVID-19 (2019-nCoV) Data Repository](https://github.com/CSSEGISandData/COVID-19). The same data sets are used to constantly update this visual dashboard: https://coronavirus.jhu.edu/map.html

**IMPORTANT DISCLAIMER:**
Although I do believe that these data and visualization help us to understand the pandemic, they are nonetheless flawed! When we want to make sense of positive test results (i.e., total cases and in the long run then mortality rates and recovering processes), we need to know how many tests were conducted. For example, Italy has an unusually high mortality rate at this moment (~8%). However, Italy ran only ~150k tests, thus the real rate of infections is probably much higher, implying a lower mortality rate. South Korea, in contrast, has tested more than 270,000 people, which amounts to more than 5200 tests per million inhabitants — more than any other country! A high diagnostic capacity at scale is hence key to epidemic control as it provides us with precise estimates and growth rate predictions. For more information, see [this article](https://www.sciencemag.org/news/2020/03/coronavirus-cases-have-dropped-sharply-south-korea-whats-secret-its-success?fbclid=IwAR3BnhqQMxCdu8-fQelEkWIDQn-j9UASV773Xl-WbIy8l7M5ZVSQpHFgkL8) in Science. For this reasons, I believe that the value of these visualization lies not in the actual numbers, but in the patterns that emerge by comparing all four curves. 

### Will these figures be updated?

Yes, I will update these figures every morning. The last update was made on `r Sys.time()`. The data of the John Hopkins University, however, is always updated at 23:59. What you see is hence the situation on `r Sys.Date()-1` at 23:59. 

## The analyses

The following code downloads the data sets and transforms them directly to be ready for the visualizations. Overall, the data is already in a very tidy format. As I am not focusing on provinces, we only summarize the cases across countries and dates (NOT provinces). That said, I distinguish China and Hong Kong due to their different timelines in responding to the virus outbreak. 

```{r, message = F, warning = F}
library(tidyverse)
library(readr)

# Function to crawl data from the github repository
get_data <- function(url) {
  read_csv(url(url)) %>%
  select(-Lat, -Long) %>%
  rename(province = "Province/State",
         country = "Country/Region") %>%
  gather(date, type, -country, -province) %>%
  mutate(type = as.numeric(type),
         date = lubridate::mdy(date)) %>% 
  tbl_df
}

# Get total cases per country/region
cases <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv" 

# Get number of deaths per country/region
deaths <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv" 

# Get number of recovered cases per country/region
recovered <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Recovered.csv" 

# Prepare data
d_cases <- get_data(cases) %>%
  rename(cases = type)
d_deaths <- get_data(deaths) %>%
  rename(deaths = type)
d_recovered <- get_data(recovered) %>%
  rename(recovered = type)

# Join data sets by country, province, and date
d <- d_cases %>%
  left_join(d_deaths) %>%
  left_join(d_recovered) 

# Distinguish Hong Kong and China
d$country[d$province=="Hong Kong"] <- "Hong Kong"

# Prepare data for visualization
data <- d %>%
  group_by(country, date) %>%
  summarize(cases = sum(cases),
            deaths = sum(deaths),
            recovered = sum(recovered)) %>%
  filter(country != "") %>%
  mutate(active = cases - (deaths + recovered)) %>%
  magrittr::set_colnames(c("country", "date", "total cases", "deaths", "recovered", "active")) %>%
  ungroup
```

### 1. Analyzing China, South Korea, Italy, and Germany

In a first step, I am comparing China, South Korea, Italy, and Germany. Why these four countries? These four countries are at different stages during the corona pandemic. China was the first to experience the outbreak and they have almost contained the spreading of the virus by now. South Korea is close to containing the virus. Italy is experiencing the second worst pandemic after China and drastic measures have been taken. The virus has reached Germany considerably later, but the growth rate is very steep. By comparing these countries, we can learn a lot about "typical" growth rates and patterns. 

```{r,results = 'asis'}
# Current cases 
table <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country == "China" | country == "Germany" | 
         country == "Italy" | country == "Korea, South") %>%
  arrange(desc(`total cases`))
papaja::apa_table(table, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```  

```{r examples, fig.height=7, fig.width=8}
# Example plot for China, Germany, Italy and South Korea
data %>%
  filter(`total cases` >= 1) %>%
  filter(country == "China" | country == "Germany" | 
         country == "Italy" | country == "Korea, South") %>%
  mutate(country = factor(country, 
                          levels = c("China", "Korea, South", 
                                     "Italy", "Germany"))) %>%
  gather(key, value, -country, -date) %>%
  ggplot(aes(x = date, y = value, color = key)) +
  geom_line() +
  scale_color_brewer(palette = "Set2") +
  facet_wrap(~country, scale = "free_y") +
  theme_bw() +
  labs(x = "date", y = "cases", color = "")

ggsave("figures/plot_1.png",
       width = 8,
       height = 7)
```

CHINA (upper left): The number of *total* confirmed cases is still rising (pink), but only very slowly and almost comes to a halt. But, the number of *active* cases (green) is declining steeply and at the same time the number of *recovered* cases (blue) is increasing a lot, slowly approximating the number of *total* cases. This is how it should look like as this pattern shows that measures are working and the spreading of the virus is stopping. 

SOUTH KOREA (upper right): The number of *total* confirmed cases is still rising (pink), but the growth rate slowly resembles a S-curve. This is a good sign, because new infections are fewer.  The number of *active* cases (green) is hence starting to decline (since the 15th of March) and at the same time the number of *recovered* cases (blue) is starting to grow.

ITALY (lower left): the number of *total* confirmed cases is growing exponentially. More importantly, the number of *active* cases is almost equivalent to the number of *total* cases (hopefully the slight deviation from the exponential trend will prove to be real). The number of *recovered* cases sadly equals the number of *deaths*. So far, we do not see implications of the drastic measures taken by the Italian government. Yet, recent analyses of the number of *total* cases suggest the curve is slowly declining (which would be a sign of hope!).

GERMANY (lower right): The number of *total* confirmed cases likewise grows exponentially. Again (see Italy), the number of *active* cases is practically equivalent to the number of *total* cases (so far, luckily only few *deaths*, but also only few recovered).

**Conclusion:** As long as we do not see signs that the curves approach a similar pattern as in China, the virus is still spreading uncontrollably. 


### 2. Development in countries that reacted fast

Although the number of cases in these countries is small (a good thing!), we should look at the distributions of countries that reacted faster (e.g., Hong Kong, Taiwan, Singapore).

```{r,results = 'asis'}
# Current cases 
table2 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country == "Hong Kong" | 
         country == "Taiwan*" | 
         country == "Singapore") %>%
  arrange(desc(`total cases`))
papaja::apa_table(table2, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```  

```{r hongkong, fig.height=3, fig.width = 8}
# Plot for Hong Kong, Taiwan, and Singapore 
data %>%
  filter(`total cases` >= 1) %>%
  filter(country == "Hong Kong" | 
         country == "Taiwan*" | 
         country == "Singapore") %>%
  gather(key, value, -country, -date) %>%
  ggplot(aes(x = date, y = value, color = key)) +
  geom_line() +
  facet_wrap(~country, scales = "free_y") +
  theme_bw() +
  scale_color_brewer(palette = "Set2") +
  labs(x = "date", y = "cases", color = "")

ggsave("figures/plot_2.png",
       width = 8,
       height = 3)
```

We clearly see that the number of *active* cases is declining earlier. At the same time the number of *total* cases increases, but not exponentially!

### 3. Worldwide developments

#### The situation in Europe

A comparative plot of all countries with more than 500 confirmed total cases.

```{r, results = 'asis'}
europe <- c("Italy", "Spain", "Germany", "France", 
            "Switzerland", "United Kingdom", "Netherlands", 
            "Norway", "Austria", "Sweden", "Belgium", "Denmark")

# Current cases 
table3 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country %in% europe) %>%
  arrange(desc(`total cases`))
papaja::apa_table(table3, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
``` 

```{r europe, fig.height=6, fig.width = 10}
data %>%
  filter(`total cases` >= 1) %>%
  filter(country %in% europe) %>%
  gather(key, value, -country, -date) %>%
  ggplot(aes(x = date, y = value, color = key)) +
  geom_line() +
  facet_wrap(~country, scales = "free_y") +
  theme_bw() +
  scale_color_brewer(palette = "Set2") +
  labs(x = "date", y = "cases", color = "")

ggsave("figures/plot_3.png",
       width = 10,
       height = 6)
```

#### USA and Canada

```{r,results = 'asis'}
# Current cases 
table4 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country == "US" | country == "Canada") %>%
  arrange(desc(`total cases`))
papaja::apa_table(table4, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```  

```{r northamerica, fig.height=3, fig.width = 7}
data %>%
  filter(`total cases` >= 1) %>%
  filter(country == "US" | country == "Canada") %>%
  gather(key, value, -country, -date) %>%
  ggplot(aes(x = date, y = value, color = key)) +
  geom_line() +
  facet_wrap(~country, scales = "free_y") +
  theme_bw() +
  scale_color_brewer(palette = "Set2") +
  labs(x = "date", y = "cases", color = "")

ggsave("figures/plot_4.png",
       width = 7,
       height = 3)
```

#### Middle East

```{r,results = 'asis'}
# Current cases 
table5 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country %in% c("Iran", "Qatar", 
                        "Israel", "Pakistan")) %>%
  arrange(desc(`total cases`))
papaja::apa_table(table5, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```  

```{r middleeast, fig.height=7, fig.width = 8}
data %>%
  filter(country %in% c("Iran", "Qatar", 
                        "Israel", "Pakistan")) %>%
  gather(key, value, -country, -date) %>%
  ggplot(aes(x = date, y = value, color = key)) +
  geom_line() +
  facet_wrap(~country, scales = "free_y") +
  theme_bw() +
  scale_color_brewer(palette = "Set2") +
  labs(x = "date", y = "cases", color = "")

ggsave("figures/plot_5.png",
       width = 8,
       height = 7)
```

#### Asia

Plot for Asian countries with > 500 cases.

```{r,results = 'asis'}
# Current cases 
table6 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country %in% c("China", "Korea, South", "Japan", "Malaysia")) %>%
  arrange(desc(`total cases`))
papaja::apa_table(table6, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```

```{r asia, fig.height=7, fig.width = 8}
data %>%
  filter(country %in% c("China", "Korea, South", "Japan", "Malaysia")) %>%
  gather(key, value, -country, -date) %>%
  ggplot(aes(x = date, y = value, color = key)) +
  geom_line() +
  facet_wrap(~country, scales = "free_y") +
  theme_bw() +
  scale_color_brewer(palette = "Set2") +
  labs(x = "date", y = "cases", color = "")

ggsave("figures/plot_6.png",
       width = 8,
       height = 7)
```


Let me know if I should include any other countries in on this page. 
