---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "figures/",
  out.width = "100%",
  fig.retina = 2,
  echo = FALSE
)
```


# Visualizing the Corona (COVID-19) pandemic

## Basic idea

I generally think that the news do a good job in describing and visualizing the corona pandemic. However, there are two things that I believe are quite problematic and that I am missing the any news coverage on the pandemic. 

1. Visualizations of the growth curves often log-transform the x-axis instead of showing the actual *exponential* growth.
2. Visualizations almost never plot *total* cases against a) deaths, b) recovered, AND c) active cases. 

With regard to the former, I recently wrote [this blogpost](http://philippmasur.de/blog/2020/03/13/understanding-exponential-growth-the-corona-pandemic/) that explains why exponential growth is so hard to grasp. With regard to the latter, I believe that visualizations would become a bit more informative if we plot total cases, active cases, recovered cases, and deaths. If we plot all *four* curves simultaneously, patterns emerge that might tell us something about the "phase"" that a country is it. We may also see whether a country succeeds in stopping the infections (BUT: see disclaimer further below!). In what follows, I am using the data on worldwide total infections, death rates, and number of recoveries to produce these (potentially) a bit more insightful visualizations. 

#### Where does the data come frome?

The analyses and visualizations are based on the data provided by the John Hopkins University in the [Official 2019 Novel Coronavirus COVID-19 (2019-nCoV) Data Repository](https://github.com/CSSEGISandData/COVID-19). The same data sets are used to constantly update this visual dashboard: https://coronavirus.jhu.edu/map.html

**IMPORTANT DISCLAIMER:**
Although I do believe that these data help us to understand the pandemic, they are nonetheless very imprecise. When we want to make sense of positive test results (i.e., total cases and in the long run mortality rates and recovering processes), we need to know how many tests were conducted. For example, Italy has an unusually high mortality rate at this moment (~8%). However, Italy ran only ~150k tests.  The real rate of infections is probably much higher, implying a lower mortality rate. South Korea, in contrast, has tested > 270,000 people, which amounts to more than 5,200 tests per million inhabitants — more tests than any other country! A high diagnostic capacity at scale is hence key to epidemic control as it provides us with precise estimates and growth rate predictions (for more information on this, see [this article](https://www.sciencemag.org/news/2020/03/coronavirus-cases-have-dropped-sharply-south-korea-whats-secret-its-success?fbclid=IwAR3BnhqQMxCdu8-fQelEkWIDQn-j9UASV773Xl-WbIy8l7M5ZVSQpHFgkL8) in Science). For this reasons, I believe that the value of these visualization lies NOT in comparing the actual numbers, but in the understanding the patterns that emerge by comparing all four curves. 

Furthermore, I would like to emphasize that I am not an expert on epidemology or virus outbreaks and I am not working in the health sector. On this page, I am only visualizing the data by the John Hopkins University in a different way that most news or other outlets do. Reliance on the these visualizations for medical guidance or use of these visualization in commerce is strictly prohibited.

#### Will these figures be updated?

Yes, I will update these figures every morning. The last update was made on `r Sys.time()`. The data of the John Hopkins University, however, are always updated at 23:59. What you see is hence the situation on `r Sys.Date()-1` at 23:59:00. Also bear in mind that the reporting of cases is somewhat delayed so that it is very likely that the actual numbers are higher.

## Visualizations

Note: I am not focusing on provinces, I hence only summarize across countries and dates. That said, I distinguish China and Hong Kong due to their different timelines in responding to the virus outbreak. 

If you are interested in the R code, please see the [README.rmd](https://github.com/masurp/corona_analyses/blob/master/README.rmd). 

```{r, message = F, warning = F}
library(tidyverse)
library(readr)

# Function to crawl data from the github repository
get_data <- function(url) {
  read_csv(url(url)) %>%
  select(-Lat, -Long) %>%
  rename(province = "Province/State",
         country = "Country/Region") %>%
  gather(date, type, -country, -province) %>%
  mutate(type = as.numeric(type),
         date = lubridate::mdy(date)) %>% 
  tbl_df
}

# Function to plot figures
plot_graph <- function(data, size = .75){
  ggplot(data, aes(x = date, y = value, color = key)) +
  geom_line(size = size) +
  scale_color_brewer(palette = "Dark2") +
  facet_wrap(~country, scale = "free_y") +
  theme_linedraw() +
  labs(x = "date", y = "cases", color = "")
}

# Get total cases per country/region
cases <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv" 

# Get number of deaths per country/region
deaths <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv" 

# Get number of recovered cases per country/region
recovered <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Recovered.csv" 

# Prepare data
d_cases <- get_data(cases) %>%
  rename(cases = type)
d_deaths <- get_data(deaths) %>%
  rename(deaths = type)
d_recovered <- get_data(recovered) %>%
  rename(recovered = type)

# Join data sets by country, province, and date
d <- d_cases %>%
  left_join(d_deaths) %>%
  left_join(d_recovered) 

# Distinguish Hong Kong and China
d$country[d$province=="Hong Kong"] <- "Hong Kong"

# Prepare data for visualization
data <- d %>%
  group_by(country, date) %>%
  summarize(cases = sum(cases),
            deaths = sum(deaths),
            recovered = sum(recovered)) %>%
  filter(country != "") %>%
  mutate(active = cases - (deaths + recovered)) %>%
  magrittr::set_colnames(c("country", "date", "total cases", "deaths", "recovered", "active")) %>%
  ungroup
```

### 1. Analyzing China, South Korea, Italy, and Germany

In a first step, I am comparing China, South Korea, Italy, and Germany. Why these countries? These four countries are at very different stages. China was the first to experience the outbreak and they have almost contained the spreading of the virus by now. South Korea is starting to control the virus outbreak. Italy is currently experiencing one of the worst outbreaks and drastic measures have been taken in the last weeks. The virus has reached Germany considerably later, but the growth rate is very steep at this moment. 

```{r example,results = 'asis'}
# Current cases 
table <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country == "China" | country == "Germany" | 
         country == "Italy" | country == "Korea, South") %>%
  arrange(desc(`total cases`))
papaja::apa_table(table, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```  

```{r, fig.height=6, fig.width=7}
# Example plot for China, Germany, Italy and South Korea
data %>%
  filter(`total cases` >= 1) %>%
  filter(country == "China" | country == "Germany" | 
         country == "Italy" | country == "Korea, South") %>%
  mutate(country = factor(country, 
                          levels = c("China", "Korea, South", 
                                     "Italy", "Germany"))) %>%
  gather(key, value, -country, -date) %>%
  plot_graph

ggsave("figures/plot_1.png",
       width = 7,
       height = 6)
```

CHINA (upper left): The number of *total* confirmed cases is still rising (pink), but only very slowly and almost comes to a halt. The number of *active* cases (green) is declining steeply and at the same time the number of *recovered* cases (blue) is increasing a lot, slowly approximating the number of *total* cases. This is how it should look like as this pattern shows that measures are working and the spreading of the virus is stopping. 

SOUTH KOREA (upper right): The number of *total* confirmed cases is still rising (pink), but the growth rate slowly resembles an S-curve. This is a good sign, because new infections are fewer. The number of *active* cases (green) is hence starting to decline (although it recently started to grow again) and at the same time the number of *recovered* cases (blue) is starting to grow.

ITALY (lower left): The number of *total* confirmed cases is still growing exponentially. More importantly, the number of *active* cases is almost equivalent to the number of *total* cases. The number of *recovered* cases sadly equals the number of *deaths*. So far, we do not see implications of the drastic measures taken by the Italian government.

GERMANY (lower right): The number of *total* confirmed cases likewise grows exponentially. Again, the number of *active* cases is practically equivalent to the number of *total* cases. 


### 2. Worldwide developments

#### Europe

Countries with more than ~500 confirmed total cases.

```{r europe, results = 'asis', comment = ""}
europe <- c("Italy", "Spain", "Germany", "France", 
            "Switzerland", "United Kingdom", "Netherlands", 
            "Norway", "Austria", "Sweden", "Belgium", "Denmark",
            "Portugal", "Czechia", "Ireland", "Turkey")

# Current cases 
table3 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country %in% europe) %>%
  arrange(desc(`total cases`))

# Plot table
papaja::apa_table(table3, 
                  format = "html", 
                  digits = 0, 
                  align = c("l", "c", rep("r", 4)))
``` 

```{r europe_plot, fig.height=5.75, fig.width = 9}
data %>%
  filter(`total cases` >= 1) %>%
  filter(country %in% europe) %>%
  gather(key, value, -country, -date) %>%
  plot_graph(size = .6) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggsave("figures/plot_3.png",
       width = 9,
       height = 5.75)
```

#### North, Middle and South America

```{r,results = 'asis'}
america <- c("US", "Canada", "Brazil")

# Current cases 
table4 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country %in% america) %>%
  arrange(desc(`total cases`))
papaja::apa_table(table4, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```  

```{r northamerica, fig.height=3, fig.width = 9}
data %>%
  filter(`total cases` >= 1) %>%
  filter(country %in% america) %>%
  gather(key, value, -country, -date) %>%
  plot_graph

ggsave("figures/plot_4.png",
       width = 9,
       height = 3)
```

#### Middle East

Countries in the middle east with > 400 cases.

```{r,results = 'asis'}
# Current cases 
table5 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country %in% c("Iran", "Qatar", 
                        "Israel", "Pakistan")) %>%
  arrange(desc(`total cases`))
papaja::apa_table(table5, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```  

```{r middleeast, fig.height=7, fig.width = 8}
data %>%
  filter(country %in% c("Iran", "Qatar", 
                        "Israel", "Pakistan")) %>%
  gather(key, value, -country, -date) %>%
  plot_graph

ggsave("figures/plot_5.png",
       width = 8,
       height = 7)
```

#### Asia

Asian countries with > 500 cases.

```{r,results = 'asis'}
# Current cases 
table6 <- data %>%
  filter(date == as.character(Sys.Date()-1)) %>%
  filter(country %in% c("China", "Korea, South", "Japan", "Malaysia")) %>%
  arrange(desc(`total cases`))
papaja::apa_table(table6, format = "html", digits = 0, align = c("l", "c", rep("r", 4)))
```

```{r asia, fig.height=7, fig.width = 8}
data %>%
  filter(country %in% c("China", "Korea, South", "Japan", "Malaysia")) %>%
  gather(key, value, -country, -date) %>%
  plot_graph

ggsave("figures/plot_6.png",
       width = 8,
       height = 7)
```



